
\addcontentsline{toc}{chapter}{Bibliography}

\begin{thebibliography}{99}

% Neural Style Transfer
\bibitem{bib:gatys2016}
L.~A. Gatys, A.~S. Ecker, and M.~Bethge,
\newblock ``Image Style Transfer Using Convolutional Neural Networks,''
\newblock {\em Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, pp. 2414--2423, 2016.

% GANs
\bibitem{bib:goodfellow2014}
I.~Goodfellow, J.~Pouget-Abadie, M.~Mirza, B.~Xu, D.~Warde-Farley, S.~Ozair, A.~Courville, and Y.~Bengio,
\newblock ``Generative Adversarial Nets,''
\newblock {\em Advances in Neural Information Processing Systems (NeurIPS)}, vol. 27, pp. 2672--2680, 2014.

% CycleGAN
\bibitem{bib:cyclegan}
J.-Y. Zhu, T.~Park, P.~Isola, and A.~A. Efros,
\newblock ``Unpaired Image-to-Image Translation using Cycle-Consistent Adversarial Networks,''
\newblock {\em Proceedings of the IEEE International Conference on Computer Vision (ICCV)}, pp. 2223--2232, 2017.

% Pix2Pix
\bibitem{bib:pix2pix}
P.~Isola, J.-Y. Zhu, T.~Zhou, and A.~A. Efros,
\newblock ``Image-to-Image Translation with Conditional Adversarial Networks,''
\newblock {\em Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, pp. 1125--1134, 2017.

% AdaIN
\bibitem{bib:adain}
X.~Huang and S.~Belongie,
\newblock ``Arbitrary Style Transfer in Real-time with Adaptive Instance Normalization,''
\newblock {\em Proceedings of the IEEE International Conference on Computer Vision (ICCV)}, pp. 1501--1510, 2017.

% Perceptual Losses
\bibitem{bib:johnson2016}
J.~Johnson, A.~Alahi, and L.~Fei-Fei,
\newblock ``Perceptual Losses for Real-Time Style Transfer and Super-Resolution,''
\newblock {\em Proceedings of the European Conference on Computer Vision (ECCV)}, pp. 694--711, 2016.

% CartoonGAN
\bibitem{bib:cartoongan}
Y.~Chen, Y.-K. Lai, and Y.-J. Liu,
\newblock ``CartoonGAN: Generative Adversarial Networks for Photo Cartoonization,''
\newblock {\em Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, pp. 9465--9474, 2018.

% AnimeGAN
\bibitem{bib:animegan}
J.~Chen, G.~Liu, and X.~Chen,
\newblock ``AnimeGAN: A Novel Lightweight GAN for Photo Animation,''
\newblock {\em International Symposium on Intelligence Computation and Applications}, pp. 242--256, 2019.

% White-box Cartoon
\bibitem{bib:whitebox}
X.~Wang and J.~Yu,
\newblock ``Learning to Cartoonize Using White-box Cartoon Representations,''
\newblock {\em Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, pp. 8090--8099, 2020.

% VGG Network
\bibitem{bib:vgg}
K.~Simonyan and A.~Zisserman,
\newblock ``Very Deep Convolutional Networks for Large-Scale Image Recognition,''
\newblock {\em International Conference on Learning Representations (ICLR)}, 2015.

% ResNet
\bibitem{bib:resnet}
K.~He, X.~Zhang, S.~Ren, and J.~Sun,
\newblock ``Deep Residual Learning for Image Recognition,''
\newblock {\em Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, pp. 770--778, 2016.

% Instance Normalization
\bibitem{bib:instancenorm}
D.~Ulyanov, A.~Vedaldi, and V.~Lempitsky,
\newblock ``Instance Normalization: The Missing Ingredient for Fast Stylization,''
\newblock {\em arXiv preprint arXiv:1607.08022}, 2016.

\bibitem{bib:ulyanov2016}
D.~Ulyanov, A.~Vedaldi, and V.~Lempitsky,
\newblock ``Instance Normalization: The Missing Ingredient for Fast Stylization,''
\newblock {\em arXiv preprint arXiv:1607.08022}, 2016.

% Universal Style Transfer
\bibitem{bib:li2017}
Y.~Li, C.~Fang, J.~Yang, Z.~Wang, X.~Lu, and M.-H. Yang,
\newblock ``Universal Style Transfer via Feature Transforms,''
\newblock {\em Advances in Neural Information Processing Systems (NeurIPS)}, vol. 30, 2017.

% Real-Time Arbitrary Stylization
\bibitem{bib:ghiasi2017}
G.~Ghiasi, H.~Lee, M.~Kudlur, V.~Dumoulin, and J.~Shlens,
\newblock ``Exploring the Structure of a Real-Time, Arbitrary Neural Artistic Stylization Network,''
\newblock {\em Proceedings of the British Machine Vision Conference (BMVC)}, 2017.

% Deep Photo Style Transfer
\bibitem{bib:luan2017}
F.~Luan, S.~Paris, E.~Shechtman, and K.~Bala,
\newblock ``Deep Photo Style Transfer,''
\newblock {\em Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, pp. 4990--4998, 2017.

% SSIT
\bibitem{bib:ssit}
C.~Oh and S.~Gonsalves,
\newblock ``SSIT: Photogenic Guided Image-to-Image Translation With Single Encoder,''
\newblock {\em IEEE Access}, vol. 11, pp. 44089--44100, 2023.

% Digital Art Style Transfer
\bibitem{bib:zhang2023}
W.~Zhang, L.~Chen, and Y.~Liu,
\newblock ``Applying Deep Learning for Style Transfer in Digital Art,''
\newblock {\em Journal of Computational Design and Engineering}, vol. 10, no. 4, pp. 1523--1535, 2023.

% Blending Art and Intelligence
\bibitem{bib:scott2024}
J.~Scott, R.~Martinez, and A.~Kumar,
\newblock ``Blending Art and Intelligence: Advances in Neural Style Transfer and Image Synthesis,''
\newblock {\em ACM Computing Surveys}, vol. 56, no. 3, pp. 1--35, 2024.

% Generative AI Art
\bibitem{bib:choi2024}
S.~Choi,
\newblock ``Generative AI Art Exploration and Image Generation Fine-Tuning Techniques,''
\newblock Master's thesis, University of California, 2024.

% LSGAN
\bibitem{bib:lsgan}
X.~Mao, Q.~Li, H.~Xie, R.~Y. Lau, Z.~Wang, and S.~P. Smolley,
\newblock ``Least Squares Generative Adversarial Networks,''
\newblock {\em Proceedings of the IEEE International Conference on Computer Vision (ICCV)}, pp. 2794--2802, 2017.

% NST Survey
\bibitem{bib:nst_survey}
Y.~Jing, Y.~Yang, Z.~Feng, J.~Ye, Y.~Yu, and M.~Song,
\newblock ``Neural Style Transfer: A Review,''
\newblock {\em IEEE Transactions on Visualization and Computer Graphics}, vol. 26, no. 11, pp. 3365--3385, 2020.

% Video Style Transfer
\bibitem{bib:video_style}
M.~Ruder, A.~Dosovitskiy, and T.~Brox,
\newblock ``Artistic Style Transfer for Videos,''
\newblock {\em German Conference on Pattern Recognition (GCPR)}, pp. 26--36, 2016.

% FID Score
\bibitem{bib:fid}
M.~Heusel, H.~Ramsauer, T.~Unterthiner, B.~Nessler, and S.~Hochreiter,
\newblock ``GANs Trained by a Two Time-Scale Update Rule Converge to a Local Nash Equilibrium,''
\newblock {\em Advances in Neural Information Processing Systems (NeurIPS)}, vol. 30, pp. 6626--6637, 2017.

% PyTorch
\bibitem{bib:pytorch}
A.~Paszke, S.~Gross, F.~Massa, A.~Lerer, J.~Bradbury, G.~Chanan, T.~Killeen, Z.~Lin, N.~Gimelshein, L.~Antiga, et al.,
\newblock ``PyTorch: An Imperative Style, High-Performance Deep Learning Library,''
\newblock {\em Advances in Neural Information Processing Systems (NeurIPS)}, vol. 32, pp. 8026--8037, 2019.

% OpenCV
\bibitem{bib:opencv}
G.~Bradski,
\newblock ``The OpenCV Library,''
\newblock {\em Dr. Dobb's Journal of Software Tools}, 2000.

% Adam Optimizer
\bibitem{bib:adam}
D.~P. Kingma and J.~Ba,
\newblock ``Adam: A Method for Stochastic Optimization,''
\newblock {\em International Conference on Learning Representations (ICLR)}, 2015.

% Attention Mechanisms
\bibitem{bib:attention}
D.~Park, S.~Yoo, H.~Paik, and G.~Kim,
\newblock ``Attention-based Multi-modal Recurrent Neural Network for Style-aware Visual Story Generation,''
\newblock {\em Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition Workshops}, pp. 71--78, 2019.

\end{thebibliography}

